---
title: "Hw4"
author: "Eshaan Singh"
date: "2025-02-15"
output: pdf_document
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(ggplot2)
library(tidyverse)
library(mosaic)
```

Link to github: https://github.com/eshaan-dev/SDS315/tree/Homework-4

# Question 1

```{r}

bootstrap_sim <- do(100000)* mean(rbinom(2021, 1, 0.024))
ggplot(bootstrap_sim) + geom_histogram(aes(x = mean),bins = 30, color = "black", fill = "lightblue")
p_value <- mean(bootstrap_sim >= 70/2021)
print(round(p_value,3))
```

The null hypothesis being tested is that if there is a probability of 2.4% trades being flagged, then there is something illegal going on and there should be an investigation into the company. 

To measure against the null hypothesis, there will be the monte-carlo simulation with a re sampling that calculates the mean of each sample. This would mean that there would be a proportion of how many iterations exceed the desired ratio and that can be filtered out using the p value constraint. 
As the p value is around `0.002`, this is a fairly small threshold and therefore can say that the null hypothesis is rejected. This would support the alternative hypothesis that their trading activity is unusual and indicates a possibility of insider trading. 

\newpage
# Question 2

```{r}

bootstrap_health <- do(100000) * sum(rbinom(1500, 1, 0.03))


ggplot(bootstrap_health, aes(x = sum)) + 
  geom_histogram(bins = 30, color = "black", fill = "lightblue3") +
  labs(title = "Distribution of Simulated Violations",
       x = "Number of Violations", y = "Frequency")

p_value <- mean(bootstrap_health$sum >= 50)
print(round(p_value,3))

```

In this case, the null hypothesis is to check if there is a standard 3% citation baseline for being flagged due to random chance.

To measure against the null hypothesis, there will be the monte-carlo simulation with a re sampling that calculates the mean of each sample. This would mean that there would be a proportion of how many iterations exceed the desired ratio and that can be filtered out using the p value constraint.

The P value differs greatly in this case as it measure to be `0.244`. This would mean that there is a 24.4% chance that there can be 8 or more violations due to chance. Since this probability is relatively high, there is not enough evidence to conclude that there's a high violation rate. In this case, the null hypothesis is accepted. 

\newpage
# Question 3

```{r}

observed <- c(85, 56, 59, 27, 13)

expected <- c(0.30, 0.25, 0.20, 0.15, 0.10) * 240

chi_squared_statistic = function(observed, expected) {
  sum((observed - expected)^2 / expected)
}

chi_res <- round(chi_squared_statistic(observed,expected),3)


chi_data <- chisq.test(observed, p = expected / sum(expected))
print(round(chi_data$p.value,3))
```

The null hypothesis in this case would be that the judge’s selection process does not systematically change the representation of its group.

Since there is a multinomial setting, it would be best to use a chi-squared test to determine statistical significance. Then using the chisq function, there can be a detailed p value that is used to back up the claim. 

After running the code, the chi-squared result was 12.426. This would mean that there is quite the difference between the observed and expected values. This difference would suggest there should be a rejection of the null. Furthermore, this is supported by the low p value of `0.014`. In conclusion, the judge’s selection process does systematically change the representation of its group.

\newpage
# Question 4

## Part A
```{r}
brown_sentences <- readLines("brown_sentences.txt")
letter_frequencies <- read.csv("letter_frequencies.csv")


letter_frequencies$Probability = letter_frequencies$Probability / sum(letter_frequencies$Probability)


calculate_chi_squared = function(sentence, freq_table) {
  

  clean_sentence = gsub("[^A-Za-z]", "", sentence)
  clean_sentence = toupper(clean_sentence)
  

  observed_counts = table(factor(strsplit(clean_sentence, "")[[1]], levels = freq_table$Letter))
  

  total_letters = sum(observed_counts)
  expected_counts = total_letters * freq_table$Probability
  

  chi_squared_stat = sum((observed_counts - expected_counts)^2 / expected_counts, na.rm=TRUE)
  
  return(chi_squared_stat)
}

chi_squared_values = sapply(brown_sentences, calculate_chi_squared, freq_table = letter_frequencies)

hist(chi_squared_values, main="Chi-Squared Distribution of Letter Frequencies")
print(round(mean(chi_squared_values),3))

```
Part A creates a null hypothesis to test the LLM watermark on. It seems like the distribution is centered around a mean of around 27. This information can be useful in Part B.


## Part B

```{r}

sentences <- c(
  "She opened the book and started to read the first chapter, eagerly anticipating what might come next.",
  "Despite the heavy rain, they decided to go for a long walk in the park, crossing the main avenue by the fountain in the center.",
  "The museum’s new exhibit features ancient artifacts from various civilizations around the world.",
  "He carefully examined the document, looking for any clues that might help solve the mystery.",
  "The students gathered in the auditorium to listen to the guest speaker’s inspiring lecture.",
  "Feeling vexed after an arduous and zany day at work, she hoped for a peaceful and quiet evening at home, cozying up after a quick dinner with some TV, or maybe a book on her upcoming visit to Auckland.",
  "The chef demonstrated how to prepare a delicious meal using only locally sourced ingredients, focusing mainly on some excellent dinner recipes from Spain.",
  "They watched the sunset from the hilltop, marveling at the beautiful array of colors in the sky.",
  "The committee reviewed the proposal and provided many points of useful feedback to improve the project’s effectiveness.",
  "Despite the challenges faced during the project, the team worked tirelessly to ensure its successful completion, resulting in a product that exceeded everyone’s expectations."
)

chi_squared_values <- sapply(sentences, calculate_chi_squared, freq_table = letter_frequencies)


p_values <- sapply(chi_squared_values, function(x) mean(chi_squared_values >= x))


results <- data.frame(Chi_Squared = round(chi_squared_values, 3), P_Value = round(p_values, 3))


print(data.frame(Index = 1:10, P_Value = results$P_Value))

```
After taking the Chi-Squared difference between the expected (null) and the outcome (10 sentences), there is a clear outlier in which the p value is less than 0.05 which is sentence 6. This sentence is the following: `Feeling vexed after an arduous and zany day at work, she hoped for a peaceful and quiet evening at home, cozying up after a quick dinner with some TV, or maybe a book on her upcoming visit to Auckland.` As it has a p-value lower than 0.05, it would mean that there is a very low chance the null hypothesis occurs, therefore hinting at the fact that it is generated by an LLM. 
